The formal verification of the ctm approach to forcing \thanks{Supported by Secyt-UNC project 33620180100465CB and Conicet.}

### 2022-08-11 Independence_CH/related.tex
Now I now where that 's' came from

We believe that our alternative formalization using the ctm approach
over Isabelle/ZF might be more appealing to set-theorists because of the
type-theoretic machinery used in Flypitch, and % might be a bit alien
since absoluteness grant us extra naturality. This last point may also
be illustrated by the treatment of ordinals; in our formalization, as
it is expected intuitively, the following are equivalent for every $x$
in a ctm $M$:
\begin{itemize}
\item $\isatt{Ord}(x)$,
\item $\isatt{ordinal}(\isatt{\#\#}M,x)$ (the relational
  definition relativized to $M$),
\item
  $M,[x] \models {\cdot}0\isatt{ is ordinal}{\cdot}$,
\item
  $M[G], [x] \models {\cdot}0\isatt{ is ordinal}{\cdot}$ for every generic $G$,
\end{itemize}
where ${\cdot}n\isatt{ is ordinal}{\cdot}$ is the code for the
appropriate first-order formula  ($0$ is a de
Bruijn index above!). In contrast, Han and van Doorn require an
injection from the ordinals of the corresponding type universe into
their encoding of a model of $\ZF$, and a further necessary injection into the Boolean
valued model using checks---this last step obviously appears in our
presentation, but the $\val$ function used to construct $M[G]$ will turn
check-names into the corresponding argument, as expected.

>>>>

We believe that our alternative formalization using the ctm approach
over Isabelle/ZF might be more appealing to set-theorists because of the
type-theoretic machinery used in Flypitch, and % might be a bit alien
since absoluteness grants us extra naturality. This last point may also
be illustrated by the treatment of ordinals; in our formalization, as
it is expected intuitively, the following are equivalent for every $x$
in a ctm $M$:
\begin{itemize}
\item $\isatt{Ord}(x)$,
\item $\isatt{ordinal}(\isatt{\#\#}M,x)$ (the relational
  definition relativized to $M$),
\item
  $M,[x] \models {\cdot}0\isatt{ is ordinal}{\cdot}$,
\item
  $M[G], [x] \models {\cdot}0\isatt{ is ordinal}{\cdot}$ for every generic $G$,
\end{itemize}
where ${\cdot}n\isatt{ is ordinal}{\cdot}$ is the code for the
appropriate first-order formula  ($0$ is a de
Bruijn index above!). In contrast, Han and van Doorn require an
injection from the ordinals of the corresponding type universe into
their encoding of a model of $\ZF$, and a further necessary injection into the Boolean
valued model using checks---this last step obviously appears in our
presentation, but the $\val$ function used to construct $M[G]$ will turn
check-names into the corresponding argument, as expected.
